---
title: "Preparing an ethnicity Read code list"
author: "Elsie Horne"
date: "23/06/2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following script loads the Read codes in rc_ethnicity.csv, and maps them to the categories used by OpenSafely:  
https://github.com/opensafely/codelist-development/issues/7#issuecomment-620206708  

```{r libraries, message=FALSE}
library(tidyverse)
```

```{r data, message=FALSE}
rc_ethnicity <- read_csv('rc_ethnicity.csv')
```

As the `vision_2` list has version 2 and version 3 Read codes, I check how these correspond to the `opensafely_2` Read codes, so I can map the categoies from version 3 to version 2.   
```{r}
rc_ethnicity %>%
  filter(!is.na(cat_vision_2), !is.na(cat_opensafely_2)) %>%
  select(read_code, cat_vision_2, cat_opensafely_2) %>%
  arrange(cat_opensafely_2) %>% 
  print(n=50)
```

These mostly correspond nicely, so cat_version_2 can be used to categorise version 2 Read codes, which are not in the OpenSafely list (OpenSafely only covers version 3), as in the below chunk. 
 
```{r}
rc_ethnicity <- rc_ethnicity %>%
  mutate(cat_final = case_when(!is.na(cat_opensafely_2) ~ cat_opensafely_2,
                               cat_vision_2 == 'Asian or Asian British: Chinese' ~ NA_character_,
                               str_detect(cat_vision_2, '^Asian') ~ 'Asian British',
                               str_detect(cat_vision_2, '^Black') ~ 'Black',
                               str_detect(cat_vision_2, '^Other') ~ 'Other',
                               str_detect(cat_vision_2, '^White') ~ 'White',
                               TRUE ~ NA_character_
                               ))
```

The only ones I have left out are those read codes in the vision_2 `Asian or Asian British: Chinese` category. I check for Read codes with 'chinese' in the Read term.

```{r}
rc_ethnicity %>% filter(str_detect(read_term_1, 
                                   regex('chinese', ignore_case = TRUE))|
                          str_detect(read_term_2, 
                                     regex('chinese', ignore_case = TRUE))) %>%
  select(read_code, read_term_1, cat_opensafely_2, cat_vision_2, cat_final) %>%
  print(n=20)
```

The Read terms for the version 2 Read codes are identical to those for the version 3 Read codes. As the version 3 ones are already categorised accoring to the cat_opensafely_2, joining on read_term should map these categories to the version 2 Read codes. Note that I remove the census year, assuming that identical read_terms from different census years will map to the same category.

```{r}
read_term_join <- rc_ethnicity %>% 
  rename(read_term = read_term_1) %>%
  # rows which have a final catego
  filter(!is.na(cat_final)) %>%
  select(read_term, cat_final) %>%
  # remove the census year
  mutate(join_var = str_trim(str_remove(read_term, '20..'), side = 'right')) %>%
  select(join_cat_final = cat_final, join_var)

rc_ethnicity <- rc_ethnicity %>% 
  rename(read_term = read_term_1) %>%
  # remove the census year
  mutate(join_var = str_trim(str_remove(read_term, '20..'), side = 'right')) %>%
  left_join(read_term_join, by = 'join_var') %>%
  mutate_at('cat_final', list(~ case_when(is.na(cat_final) & !is.na(join_cat_final) ~ join_cat_final,
                                          TRUE ~ .))) %>%
  select(-starts_with('join'))
```

Check the read_codes which have not yet been assigned a cat_final cateogry.

```{r}
rc_ethnicity %>%
  filter(is.na(cat_final), !is.na(read_term)) %>%
  select(read_term) %>%
  print(n=20)
```

I manually sort out these reamaining ones.

```{r}
rc_ethnicity <- rc_ethnicity %>%
  mutate_at('cat_final', list(~ case_when(!is.na(cat_final) ~ .,
                                          is.na(cat_final) & read_term %in% c('Mid East (excl Israeli  Iranian & Arab) - eth cat 2001 cens') ~ 'Other',
                                          is.na(cat_final) & read_term %in% c('Asian: Indian Indian Scot/Indian Brit- Scotland 2011 census',
                                                           'Bangladeshi Bangladeshi Scot or Bangladeshi Brit- Scot 2011') ~ 'Asian British',
                                          TRUE ~ NA_character_))) %>%
  filter(!is.na(cat_final)) %>%
  select(read_code, read_term, cat_final)
```

```{r}
write_csv(rc_ethnicity, path = 'rc_ethnicity_final.csv')
```

