---
title: "Preparing an ethnicity Read code list"
author: "Elsie Horne"
date: "23/06/2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following script loads `rc_ethnicity.csv`, the lists of Read codes scraped in `read_tables.R`. All Read codes are then mapped to the [categories used in the OpenSafely project](https://github.com/opensafely/codelist-development/issues/7#issuecomment-620206708)  

```{r libraries, message=FALSE}
library(tidyverse)
```

```{r data, message=FALSE}
rc_ethnicity <- read_csv('rc_ethnicity.csv') %>%
  mutate(cat_final = cat_opensafely_2)
```

View the data where the Read codes are in the OpenSafely list:  

```{r}
rc_ethnicity %>%
  filter(!is.na(cat_final)) %>%
  select(read_code, read_term_1, read_term_2, cat_final)
```

View the data where the Read codes are *not* in the OpenSafely list:  

```{r}
rc_ethnicity %>%
  filter(is.na(cat_final)) %>%
  select(read_code, read_term_1, read_term_2, cat_final)
```

As the OpenSafely Read codes are all version 3 Read codes, I map the categories (`cat_opensafely_2`) to the version 2 Read codes by joining on the Read term (`read_term`). I remove the census year from `read_term` before doing so. There are two key assumptions here:  
1. All Read terms correspond to the correct Read code (i.e. there are no errors in the tables that I scraped from the websites).  
2. The 2001 and 2011 census Read codes should be mapped to the same category if identical other than the year.  

I do this first for `read_term_1`, and will repeat for `read_term_2` if necessary.

```{r}
read_term_join <- rc_ethnicity %>% 
  rename(read_term = read_term_1) %>%
  # rows which have a final category
  filter(!is.na(cat_final)) %>%
  select(read_term, cat_final) %>%
  # remove the census year
  mutate(join_var = str_trim(str_remove(read_term, '20..'), side = 'right')) %>%
  select(join_cat_final = cat_final, join_var)

rc_ethnicity <- rc_ethnicity %>% 
  rename(read_term = read_term_1) %>%
  # remove the census year
  mutate(join_var = str_trim(str_remove(read_term, '20..'), side = 'right')) %>%
  left_join(read_term_join, by = 'join_var') %>%
  mutate_at('cat_final', list(~ case_when(is.na(cat_final) & !is.na(join_cat_final) ~ join_cat_final,
                                          TRUE ~ .))) %>%
  select(-starts_with('join'))
```

Check the read_codes which have not yet been assigned a cat_final category.

```{r}
rc_ethnicity %>%
  filter(is.na(cat_final)) %>%
  select(read_code, read_term) %>%
  print(n=20)
```

Manually sort out these reamaining ones. The missing term for Read code (XaJRB) is 'Ethnic category not stated - 2001 census'.

```{r}
rc_ethnicity <- rc_ethnicity %>%
  mutate_at('cat_final', list(~ case_when(!is.na(cat_final) ~ .,
                                          is.na(cat_final) & read_term %in% c('Mid East (excl Israeli  Iranian & Arab) - eth cat 2001 cens') ~ 'Other',
                                          is.na(cat_final) & read_term %in% c('Asian: Indian Indian Scot/Indian Brit- Scotland 2011 census',
                                                           'Bangladeshi Bangladeshi Scot or Bangladeshi Brit- Scot 2011') ~ 'Asian British',
                                          TRUE ~ NA_character_))) %>%
  filter(!is.na(cat_final)) %>%
  select(read_code, read_term, cat_final)
```

```{r}
rc_ethnicity %>%
  group_by(cat_final) %>%
  count()
```

```{r}
write_csv(rc_ethnicity, path = 'rc_ethnicity_final.csv')
```

